### 2014-11-13 coordination.tex
Fixed more typos.

In order to allow different scheduling strategies, we introduce the concept of
\emph{node priority} by assigning a priority value to every node in the program
and by introducing coordination facts that manipulate such priority values.
By default, nodes have no priority and can be picked in any order. In
our implementation, we use a FIFO approach because
older nodes tend to a higher number of unexamined facts, from which to derive
subsequent new facts.

We have two kinds of priorities: a \emph{temporary priority} and a \emph{default
priority}. A temporary priority momentarily changes the default priority $D$ of a
node, so that once the node is done, the priority will default back to $D$.
Initially, all nodes have a default priority of $0$.

@@

\begin{tightitemize}
   \item \code{set-priority(node A, float F)}: This sets the
   temporary priority of \mytt{A} to \mytt{F}. If \mytt{A} has a priority
   \mytt{F'}, we only change the priority if \mytt{F} is higher. The programmer
   can decide if priorities are to be ordered in ascending or descending order.
   \item \code{add-priority(node A, float F)}: Increases,
   temporarily, the priority of node \mytt{A} by \mytt{F}.
   \item \code{remove-priority(node A)}: Removes the temporary priority from node
   \mytt{A}.
   \item \code{schedule-next(node A)}: Changes the temporary priority of node
   \mytt{A} to be 1 unit higher than the current highest priority.
   \item \code{set-default-priority(node, float)}: Sets the default
   priority of the node.
   \item \code{stop-program(node A)}: Immediately stops the
   execution of the whole program.
\end{tightitemize}

CLM provides the sensing fact \code{priority(node A, node B, float P)} in order to sense the
priority \mytt{P} of node \mytt{B} (at node \mytt{A}).
Sensing facts are only used in the body of the rules in order
to fetch information from the runtime system.
Note that when sensing facts are used to prove new facts, they are re-derived automatically.
~\footnote{The observant reader will notice that all the coordination facts are
linear and when created in the body of a rule will be consumed. The system
creates the necessary code to re-derive them without programmer interaction.
Likewise, \mytt{set-priority} and \mytt{set-default-priority} update the
value of \mytt{priority} facts by consuming and re-deriving them but this is
done automatically by the runtime system.}

@@

\begin{tightitemize}
   \item \code{cpu-id(node A, node B, int T)}: Store at node \mytt{A} which
   thread \mytt{T}, \mytt{B} is actually running on.
   \item \code{moving(node A, node B)}: Fact available at node \mytt{A} if \mytt{B} is allowed
   to move between threads.
   \item \code{static(node A, node B)}: Fact available at node \mytt{A} if \mytt{B} is not
   allowed to move between threads.
\end{tightitemize}

>>>>

In order to allow different scheduling strategies, we introduce the concept of
\emph{node priority} by assigning a priority value to every node in the program
and by introducing coordination facts that manipulate such priority values.
By default, nodes have no priority and can be picked in any order. In
our implementation, we use a FIFO approach because
older nodes tend to have a higher number of unexamined facts, from which to derive
subsequent new facts.

We have two kinds of priorities: a \emph{temporary priority} and a \emph{default
priority}. A temporary priority momentarily changes the default priority $D$ of a
node, so that once the node is done, the priority will default back to $D$.
Initially, all nodes have a default priority of $-\infty$.

@@

\begin{tightitemize}
   \item \code{set-priority(node A, float F)}: This sets the
   temporary priority of \mytt{A} to \mytt{F}. If \mytt{A} has priority
   \mytt{F'}, we only change the priority if \mytt{F} is higher. The programmer
   can decide if priorities are to be ordered in ascending or descending order.
   \item \code{add-priority(node A, float F)}: Increases,
   temporarily, the priority of node \mytt{A} by \mytt{F}.
   \item \code{remove-priority(node A)}: Removes the temporary priority from node
   \mytt{A}.
   \item \code{schedule-next(node A)}: Changes the temporary priority of node
   \mytt{A} to be $+\infty$.
   \item \code{set-default-priority(node, float)}: Sets the default
   priority of the node.
   \item \code{stop-program(node A)}: Immediately stops the
   execution of the whole program.
\end{tightitemize}

CLM provides the sensing fact \code{priority(node A, node B, float P)} in order to sense the
priority \mytt{P} of node \mytt{B} (at node \mytt{A}).
Sensing facts are only used in the body of the rules in order
to fetch information from the runtime system.
Note that when sensing facts are used to prove new facts, they are re-derived automatically.
~\footnote{The observant reader will notice that all the coordination facts are
linear and when used in the body of a rule will be consumed. The system
creates the necessary code to re-derive them without programmer interaction.
Likewise, \mytt{set-priority} and \mytt{set-default-priority} update the
value of \mytt{priority} facts by consuming and re-deriving them but this is
done automatically by the runtime system.}

@@

\begin{tightitemize}
   \item \code{cpu-id(node A, node B, int T)}: Store at node \mytt{A} which
   thread \mytt{T}, \mytt{B} is actually running on.
   \item \code{is-moving(node A, node B)}: Fact available at node \mytt{A} if \mytt{B} is allowed
   to move between threads.
   \item \code{is-static(node A, node B)}: Fact available at node \mytt{A} if \mytt{B} is not
   allowed to move between threads.
\end{tightitemize}

### 2014-11-13 language.tex
Fix verbatim issues.
Small improvements.
Fixed more typos.

Coordinated Linear Meld (CLM) is a forward-chaining linear logic programming
language based on LM~\cite{cruz-iclp14}.
Programs consist of a set of \emph{rules} and a \emph{database of facts}.
Rules have the form \mytt{a(X), b(Y) -o c(X, Y)} and can be read as follows: if fact \mytt{a(X)}
and fact \mytt{b(Y)} exist in the database then \mytt{c(X, Y)} is added to
the database. The expression \mytt{a(X), b(Y)} is called the \emph{body} of
the rule and \mytt{c(X, Y)} is the \emph{head} of the rule.
A fact is a predicate, e.g., \mytt{a}, \mytt{b} or \mytt{c}, and its
associated tuple of values, e.g., the concrete values of \mytt{X} and
\mytt{Y}. Since CLM uses linear logic as its foundation, we distinguish between
\emph{linear} and \emph{persistent facts}. Linear facts are deleted during the
process of deriving a rule, while persistent facts are not.
Program execution starts by adding the \emph{axioms} (the initial facts of the
program) to the database. Next, rules are recursively applied and the
database is updated by adding new facts or deleting facts used during rule derivation.
When no more rules are applicable, the program terminates.

@@

\paragraph{SSSP Example}\label{sect:ssspex}
To make these ideas concrete, Fig.~\ref{code:shortest_path_program} presents
a simple program: the single source shortest path program~(SSSP). Later in the paper, we
add coordination facts to improve the execution of the program.

>>>>

Coordinated Linear Meld (CLM) is a forward-chaining linear logic programming
language based on LM~\cite{cruz-iclp14}.
Programs consist of a set of \emph{rules} and a \emph{database of facts}.
Rules have the form \mytt{a(X), b(Y) -o c(X, Y)} and can be read as follows: if fact \mytt{a(X)}
and fact \mytt{b(Y)} exist in the database then \mytt{c(X, Y)} is added to
the database. The expression \mytt{a(X), b(Y)} is called the \emph{body} of
the rule and \mytt{c(X, Y)} is the \emph{head} of the rule.
A fact is a predicate, e.g., \mytt{a}, \mytt{b} or \mytt{c}, and its
associated tuple of values, e.g., the concrete values of \mytt{X} and
\mytt{Y}. Since CLM uses linear logic as its foundation, we distinguish between
\emph{linear} and \emph{persistent facts}. Linear facts are deleted during the
process of deriving a rule, while persistent facts are not.
Program execution starts by adding the \emph{axioms} (the initial facts)
to the database. Next, rules are recursively applied and the
database is updated by adding new facts or deleting facts used during rule derivation.
When no more rules are applicable, the program terminates.

@@

\paragraph{SSSP Example}\label{sect:ssspex}
To make these ideas concrete, Fig.~\ref{code:shortest_path_program} presents
a simple program: the single source shortest path program~(SSSP). Later in the paper, we
add coordination facts to improve the execution of this program.

### 2014-11-13 related.tex
Fixed more typos.

\paragraph{Forward-Chaining Logic Programming}
This paradigm came into its own with the Datalog
language~\cite{Ullman:1990:PDK:533142}.  Traditionally used in deductive
databases, it is now being increasingly used in different fields such as
distributed networking~\cite{Loo-condie-garofalakis-p2}, sensor
nets~\cite{Chu:2007:DID:1322263.1322281} and cloud computing~\cite{alvaro:boom}.
The LM language was inspired by Meld~\cite{ashley-rollman-iclp09}, a
Datalog-like language for programming distributed ensembles of modular robots.
Meld also introduced the idea of sensing and action facts in order to sense and act
on the outside world, respectively. LCM uses sensing facts to sense the
underlying system and action facts to act on that system.

>>>>

\paragraph{Forward-Chaining Logic Programming}
This paradigm came into its own with the Datalog
language~\cite{Ullman:1990:PDK:533142}.  Traditionally used in deductive
databases, it is now being used for
distributed networking~\cite{Loo-condie-garofalakis-p2}, sensor
nets~\cite{Chu:2007:DID:1322263.1322281} and cloud computing~\cite{alvaro:boom}.
The LM language was inspired by Meld~\cite{ashley-rollman-iclp09}, a
Datalog-like language for programming distributed ensembles of modular robots.
Meld also introduced the idea of sensing and action facts in order to sense and act
on the outside world, respectively. LCM uses sensing facts to sense the
underlying system and action facts to act on that system.
